### 2016-03-04 manuscript.Rmd
fill out the YAML

---
title: "Modelling palaeoenvironmental time series and feature detection"
author:
- name: Gavin L Simpson
  affiliation: Insitute of Environmental CHange and Society, University of Regina, Regina, SK
date: March 4, 2016
output: 
  pdf_document:
    keep_tex: true
    md_extensions: +header_attributes+superscript+subscript
#    pandoc_args: [
#      "--no-tex-ligatures"
#    ]
#    template: frontiers-template.tex
# documentclass: frontiersSCNS # not yet ready for this; need to pandoc-ify the template
bibliography: references.bib
---

## Abstract

350 Words

## Introduction